Score-based tests of measurement invariance: Use in practice

### 2013-09-25 frontiers/frontiers.Rnw
Revisions to frontiers paper, simulation sections.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@265 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Results}
\readme{When referring to statistics, it would be good to use the notation from Equations 6--8.  For example, instead of ``unordered $\max\mathit{LM}$'', you would have $\max\mathit{LM}_{uo}$.}
Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibited 
similar results, with unordered  $\max\mathit{LM}$ demonstrated lower powers across all situations. ordered $\max\mathit{LM}$ and double-max exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

From these figures, one generally observe that the tests isolate 
the parameter violating measurement invariance.  Additionally, the tests 
have somewhat-higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation.

\readme{We need to compare the specific tests: the unordered test, which mimics the LRT, has lower power.}

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified. \readme{The plots have too many points; maybe we should only plot  every other diff value.  In the xyplot command, you could add the argument \verb+subset = diff %% .5 == 0+.}

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")  
@

\begin{figure}
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Loading1.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7, echo=FALSE>>=
sim1.tmp <- sim1[sim1$parms %in% c("loading")& sim1$pars %in% c("Loading1","All Loadings","All Errors"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

>>>>

\subsection{Results}
\readme{When referring to statistics, it would be good to use the notation from Equations 6--8.  For example, instead of ``unordered $\max\mathit{LM}$'', you would have $\max\mathit{LM}_{uo}$.}
Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibited 
similar results, with unordered  $\max\mathit{LM}$ demonstrating lower power across all situations. ordered $\max\mathit{LM}$ and double-max exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent. \readme{Should say that the unordered statistic is asymptotically equivalent to the LRT.}

From these figures, one generally observes that the tests isolate 
the parameter violating measurement invariance.  Additionally, the tests 
have somewhat-higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation.

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified. \readme{The plots have too many points; maybe we should only plot  every other diff value.  In the xyplot command, you could add the argument ``subset = diff \%\% .5 == 0''.}

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(lambda[11] - lambda[62]), 
               expression(psi[11] - psi[66]))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")  
@

\begin{figure}
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ test, the double-max test, and the unordered $\max\mathit{LM}$ test across three
      levels of the ordinal variable $m$ and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is $\lambda_{11}$.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7, echo=FALSE>>=
#sim1.tmp <- sim1[sim1$parms %in% c("loading")& sim1$pars %in% c("Loading1","All Loadings","All Errors"),]
#sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1,
             subset = (parms=="loading" & 
                      pars %in% c("Loading1","All Loadings","All Errors")),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey,
             strip = function(..., which.given, factor.levels){
                 if(which.given==2){
                     strip.default(which.given, factor.levels = parlabs[c(1,4:5)], ...)
                 } else {
                     strip.default(which.given, factor.levels = factor.levels, ...)
                 }
             }))
@
